New lower bounds for Schur and weak Schur numbers

### 2021-06-12 article/weakschur.tex
environnements figure
modifs

\begin{figure}[H]
\caption{Construction of the weakly sum-free coloring}
\label{SchemaWS}
\[
\begin{NiceArray}{*{13}{c}}[corners={NW,SW},hvlines,first-row,last-row,first-col]
\CodeBefore
	\cellcolor{green}{1-6}
	\cellcolor{green}{1-7}
	\cellcolor{cyan}{1-8}
	\cellcolor{green}{1-9}
	\cellcolor{cyan}{1-10}
	\cellcolor{cyan}{1-11}
	\cellcolor{cyan}{1-12}
	\cellcolor{green}{1-13}
	\cellcolor{red}{2-1}
	\cellcolor{red}{2-2}
	\cellcolor{red}{2-3}
	\cellcolor{red}{2-4}
	\cellcolor{red}{2-5}
	\cellcolor{red}{2-6}
	\cellcolor{red}{2-7}
	\cellcolor{red}{2-8}
	\cellcolor{red}{2-9}
	\cellcolor{cyan}{2-10}
	\cellcolor{cyan}{2-11}
	\cellcolor{cyan}{2-12}
	\cellcolor{green}{2-13}
	\cellcolor{yellow}{3-1}
	\cellcolor{yellow}{3-2}
	\cellcolor{yellow}{3-3}
	\cellcolor{yellow}{3-4}
	\cellcolor{yellow}{3-5}
	\cellcolor{yellow}{3-6}
	\cellcolor{yellow}{3-7}
	\cellcolor{yellow}{3-8}
	\cellcolor{yellow}{3-9}
	\cellcolor{cyan}{3-10}
	\cellcolor{cyan}{3-11}
	\cellcolor{cyan}{3-12}
	\cellcolor{green}{3-13}
	\cellcolor{yellow}{4-1}
	\cellcolor{yellow}{4-2}
	\cellcolor{yellow}{4-3}
	\cellcolor{yellow}{4-4}
	\cellcolor{yellow}{4-5}
	\cellcolor{yellow}{4-6}
	\cellcolor{yellow}{4-7}
	\cellcolor{yellow}{4-8}
	\cellcolor{yellow}{4-9}
	\cellcolor{cyan}{4-10}
	\cellcolor{cyan}{4-11}
	\cellcolor{cyan}{4-12}
	\cellcolor{green}{4-13}
	\cellcolor{red}{5-1}
	\cellcolor{red}{5-2}
	\cellcolor{red}{5-3}
	\cellcolor{red}{5-4}
	\cellcolor{red}{5-5}
	\cellcolor{red}{5-6}
	\cellcolor{red}{5-7}
	\cellcolor{red}{5-8}
	\cellcolor{red}{5-9}
	\cellcolor{cyan}{5-10}
	\cellcolor{cyan}{5-11}
	\cellcolor{cyan}{5-12}
	\cellcolor{green}{5-13}
	\cellcolor{cyan}{6-10}
	\cellcolor{cyan}{6-11}
	\cellcolor{cyan}{6-12}
	\cellcolor{green}{6-13}
\Body
	& & & & & & \Block{1-8}{\overbrace{\hphantom{---------------------------}}^{\mathcal{T}}} \\
	& & & & & & 1 & 2 & ... & r & r + 1 & ... & b - 1 & b \\
	\Block{5-1}{\mathcal{R} \left\{ \vphantom{\begin{array}{l} . \\ . \\ . \\ . \\ . \\ . \\ . \\ . \\ . \end{array}} \right.} 
	& a - l & a - l + 1 & ... & a - 1 & a & a + 1 & ... & a + r - 1 & a + r & a + r + 1 & ... & a + b - 1 & a + b \\
	& 2 a - l & ... & ... & ... & 2 a & ... & ... & ... & 2 a + r & ... & ... & ... & 2 a + b \\
	& ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... \\
	& ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... \\
	& p a - l & ... & ... & ... & p a & ... & ... & ... & p a + r & ... & ... & ... & p a + b \\
	& & & & & & & & & & \Block{1-4}{\underbrace{\hphantom{--------------}}_{\mathcal{C}}} \\
\end{NiceArray}
\]
\end{figure}

@@

\begin{sloppypar}
\(\pi\) is the projection on the first line mentioned in the intuitive explanation. The following four propositions are
calculative properties on \(\pi\) reflecting the behaviour of an element's column number in \hyperref[SchemaWS]{Figure \ref{SchemaWS}} and
that we will use later when we introduce WSF-templates.
\end{sloppypar}

@@

After defining the function related to the column number of each element in \hyperref[SchemaWS]{Figure \ref{SchemaWS}},
we introduce the function related to its line number.

@@

\begin{proposition}
Let \((b, k, a) \in (\mathbb{N}^*)^3\) and let \(f\) be a coloring associated to a \(b\)-WSF-template with width \(p\) and \(k\) colors. Let
\(c \in \mathbb{N}\) (\(c = \min (A_{k+1} \backslash [\![1, b ]\!]) - 1\) works) and assume there there exists a coloring \(g\) of
\([\![b + 1, b + c]\!]\) with \(k\) colors such that for all \(c \in [\![1, k]\!]\),

@@

\begin{theorem}
\label{Theorem3.22}
Let \((k,p) \in (\mathbb{N}^*)^2\) and \((a, n, b) \in (\mathbb{N}^*)^3\) . If there exists a SF-template with width 
\(p\) and \(k+1\) colors and a \(b\)-WSF-template with width \(a\) and \(n\) colors, then there exists \(pb\)-WSF-template 
with width \(pq\) and \((n+k)\) colors.
\end{theorem}

@@

\begin{align}
	WS(n + 1) &\geqslant  4 S(n)  +  2 \label{WS(n+1)} \\
	WS(n + 2) &\geqslant  13 S(n)  +  8 \label{WS(n+2)} \\
	WS(n + 3) &\geqslant  42 S(n)  +  24 \label{WS(n+3)} \\
	WS(n + 4) &\geqslant  132 S(n)  +  26 \label{WS(n+4)}
\end{align}

\hyperref[WS(n+1)]{(\ref{WS(n+1)})} and \hyperref[WS(n+2)]{(\ref{WS(n+2)})} were already found by Rowley, they are detailed in \cite{RowleyWS}. \hyperref[WS(n+3)]{(\ref{WS(n+3)})} is optimal 
and was found with a SAT solver. It uses the first sophistication explained in the \hyperref[ConstructionWSF]{Subsection 3.3} in order to add 
the last number in the first color. The \hyperref[WSF-templates]{corresponding template} can be found in the appendix.
As for \hyperref[WS(n+4)]{(\ref{WS(n+4)})}, it was obtained by combining an optimal SF-template with width 33 with a WSF-template with width 4.
The best template we could get with a computer search gives the inequality \(\WS(n+4) \geqslant 127 S(n) + 68\).
It was also found with the SAT solver. In order to reduce the search space, we only looked for WSF-templates of
five colors which start with a good \(\WS(4)\) partition and we assumed that the special color was the last by order of 
appearance. However, this approach most likely prevents us from finding the best WSF-templates
as we explain in the next subsection for weakly sum-free partitions. We highly suspect that there exists more efficient WSF-templates
with \(n \geqslant 5\) colors. One may try to go over a different search space using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}.
This could be the suject of a future work. Further details about the encoding as a SAT problem can be found in \cite{Heule2017}.

Like in Subsection 2.3 \textbf{LABEL NECESSAIRE}, we compute the lower bounds given by \hyperref[WS(n+1)]{(\ref{WS(n+1)})}, 
\hyperref[WS(n+2)]{(\ref{WS(n+2)})} and \hyperref[WS(n+3)]{(\ref{WS(n+3)})} for \( n \in [\![8,15]\!] \). The best lower bound
for each integer is highlighted.

@@

\begin{table}[H]
\caption{New lower bounds for \( n \in [\![8,15]\!] \)}
\label{LowerBoundsWS}
\[
\begin{array}{c}
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-4}
		\cellcolor{yellow}{2-5}
		\cellcolor{yellow}{3-2}
		\cellcolor{yellow}{4-3}
	\Body
		n & 8 & 9 & 10 & 11 \\
		4 \, S(n-1) + 2 & 6\,722 & 21\,146 & 71\,214 & 243\,794 \\
		13 \, S(n-2) + 8 & 6\,976 & 21\,848 & 68\,726 & 231\,447 \\
		42 \, S(n-3) + 24 & 6\,744 & 22\,536 & 70\,584 & 222\,036 \\
	\end{NiceArray}
	\\ \\
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-2}
		\cellcolor{yellow}{2-5}
		\cellcolor{yellow}{3-3}
		\cellcolor{yellow}{4-4}
	\Body
		n & 12 & 13 & 14 & 15 \\
		4 \, S(n-1) + 2 & 815\,314 & 2\,554\,194 & 8\,045\,162 & 27\,061\,154 \\
		13 \, S(n-2) + 8 & 792\,332 & 2\,649\,772 & 8\,301\,132 & 26\,146\,778 \\
		42 \, S(n-3) + 24 & 747\,750 & 2\,559\,840 & 8\,560\,800 &  25\,886\,224 \\
	\end{NiceArray}
\end{array}
\]
\end{table}

@@

With \( S(9) \geqslant 17\,803 \), we found a new lower bound for \(\WS(10)\) using \hyperref[WS(n+1)]{(\ref{WS(n+1)})}.
Moreover, \hyperref[WS(n+3]{(\ref{WS(n+3)})} gives new lower bounds for \(\WS(9)\) and \(WS(14)\).

@@

\qquad In this section, we first gave a new construction which can be seen as an equivalent for weakly sum-free partitions of Abott
and Hanson's construction for sum-free partitions. We then  introduced WSF-templates and generalized this construction. This
allows us to find new lower bounds and new inequalities for weak Schur numbers. One may notice the significant gap
between the former lower bounds for weak Schur numbers obtained by conducting a computer search and the new lower bounds
obtained with WSF-templates (including \hyperref[WS(n+1)]{(\ref{WS(n+1)})} and \hyperref[WS(n+2]{(\ref{WS(n+2)})}).

>>>>

\begin{figure}[H]

\label{SchemaWS}
\[
\begin{NiceArray}{*{13}{c}}[corners={NW,SW},hvlines,first-row,last-row,first-col]
\CodeBefore
	\cellcolor{green}{1-6}
	\cellcolor{green}{1-7}
	\cellcolor{cyan}{1-8}
	\cellcolor{green}{1-9}
	\cellcolor{cyan}{1-10}
	\cellcolor{cyan}{1-11}
	\cellcolor{cyan}{1-12}
	\cellcolor{green}{1-13}
	\cellcolor{red}{2-1}
	\cellcolor{red}{2-2}
	\cellcolor{red}{2-3}
	\cellcolor{red}{2-4}
	\cellcolor{red}{2-5}
	\cellcolor{red}{2-6}
	\cellcolor{red}{2-7}
	\cellcolor{red}{2-8}
	\cellcolor{red}{2-9}
	\cellcolor{cyan}{2-10}
	\cellcolor{cyan}{2-11}
	\cellcolor{cyan}{2-12}
	\cellcolor{green}{2-13}
	\cellcolor{yellow}{3-1}
	\cellcolor{yellow}{3-2}
	\cellcolor{yellow}{3-3}
	\cellcolor{yellow}{3-4}
	\cellcolor{yellow}{3-5}
	\cellcolor{yellow}{3-6}
	\cellcolor{yellow}{3-7}
	\cellcolor{yellow}{3-8}
	\cellcolor{yellow}{3-9}
	\cellcolor{cyan}{3-10}
	\cellcolor{cyan}{3-11}
	\cellcolor{cyan}{3-12}
	\cellcolor{green}{3-13}
	\cellcolor{yellow}{4-1}
	\cellcolor{yellow}{4-2}
	\cellcolor{yellow}{4-3}
	\cellcolor{yellow}{4-4}
	\cellcolor{yellow}{4-5}
	\cellcolor{yellow}{4-6}
	\cellcolor{yellow}{4-7}
	\cellcolor{yellow}{4-8}
	\cellcolor{yellow}{4-9}
	\cellcolor{cyan}{4-10}
	\cellcolor{cyan}{4-11}
	\cellcolor{cyan}{4-12}
	\cellcolor{green}{4-13}
	\cellcolor{red}{5-1}
	\cellcolor{red}{5-2}
	\cellcolor{red}{5-3}
	\cellcolor{red}{5-4}
	\cellcolor{red}{5-5}
	\cellcolor{red}{5-6}
	\cellcolor{red}{5-7}
	\cellcolor{red}{5-8}
	\cellcolor{red}{5-9}
	\cellcolor{cyan}{5-10}
	\cellcolor{cyan}{5-11}
	\cellcolor{cyan}{5-12}
	\cellcolor{green}{5-13}
	\cellcolor{cyan}{6-10}
	\cellcolor{cyan}{6-11}
	\cellcolor{cyan}{6-12}
	\cellcolor{green}{6-13}
\Body
	& & & & & & \Block{1-8}{\overbrace{\hphantom{---------------------------}}^{\mathcal{T}}} \\
	& & & & & & 1 & 2 & ... & r & r + 1 & ... & b - 1 & b \\
	\Block{5-1}{\mathcal{R} \left\{ \vphantom{\begin{array}{l} . \\ . \\ . \\ . \\ . \\ . \\ . \\ . \\ . \end{array}} \right.} 
	& a - l & a - l + 1 & ... & a - 1 & a & a + 1 & ... & a + r - 1 & a + r & a + r + 1 & ... & a + b - 1 & a + b \\
	& 2 a - l & ... & ... & ... & 2 a & ... & ... & ... & 2 a + r & ... & ... & ... & 2 a + b \\
	& ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... \\
	& ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... \\
	& p a - l & ... & ... & ... & p a & ... & ... & ... & p a + r & ... & ... & ... & p a + b \\
	& & & & & & & & & & \Block{1-4}{\underbrace{\hphantom{--------------}}_{\mathcal{C}}} \\
\end{NiceArray}
\]
\caption{Construction of the weakly sum-free coloring}
\end{figure}

@@

\begin{sloppypar}
\(\pi\) is the projection on the first line mentioned in the intuitive explanation. The following four propositions are
calculative properties on \(\pi\) reflecting the behaviour of an element's column number in Figure \ref{SchemaWS} and
that we will use later when we introduce WSF-templates.
\end{sloppypar}

@@

After defining the function related to the column number of each element in Figure \ref{SchemaWS},
we introduce the function related to its line number.

@@

\begin{proposition}
Let \((b, k, a) \in (\mathbb{N}^*)^3\) and let \(f\) be a coloring associated to a \(b\)-WSF-template with width \(p\) and \(k\) colors. Let
\(c \in \mathbb{N}\) and assume there there exists a coloring \(g\) of \([\![b + 1, b + c]\!]\) with \(k\) colors such that for all \(c \in [\![1, k]\!]\),

@@

\begin{theorem}
\label{Theorem3.22}
Let \((k,p) \in (\mathbb{N}^*)^2\) and \((a, n, b) \in (\mathbb{N}^*)^3\) . If there exists a SF-template with width 
\(p\) and \(k+1\) colors and a \(b\)-WSF-template with width \(a\) and \(n\) colors, then there exists a \(pb\)-WSF-template 
with width \(pq\) and \((n+k)\) colors.
\end{theorem}

@@

\begin{align}
	WS(n + 1) &\geqslant  4\,S(n)  +  2 \label{WS(n+1)} \\
	WS(n + 2) &\geqslant  13\,S(n)  +  8 \label{WS(n+2)} \\
	WS(n + 3) &\geqslant  42\,S(n)  +  24 \label{WS(n+3)} \\
	WS(n + 4) &\geqslant  132\,S(n)  +  26 \label{WS(n+4)}
\end{align}

(\ref{WS(n+1)}) and (\ref{WS(n+2)}) were already found by Rowley, they are detailed in \cite{RowleyWS}. (\ref{WS(n+3)}) is optimal 
and was found with a SAT solver. It uses the first sophistication explained in the \hyperref[ConstructionWSF]{Subsection 3.3} in order to add 
the last number in the first color. The \hyperref[WSF-templates]{corresponding template} can be found in the appendix.
As for (\ref{WS(n+4)}), it was obtained by combining an optimal SF-template with width 33 with a WSF-template with width 4.
The best template we could get with a computer search gives the inequality \(\WS(n+4) \geqslant 127 S(n) + 68\).
It was also found with the SAT solver. In order to reduce the search space, we only looked for WSF-templates of
five colors which start with a good \(\WS(4)\) partition and we assumed that the special color was the last by order of 
appearance. However, this approach most likely prevents us from finding the best WSF-templates
as we explain in the next subsection for weakly sum-free partitions. We highly suspect that there exists more efficient WSF-templates
with \(n \geqslant 5\) colors. One may try to go over a different search space using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}.
This could be the suject of a future work. Further details about the encoding as a SAT problem can be found in \cite{Heule2017}.

Like in Subsection 2.3 \textbf{LABEL NECESSAIRE}, we compute the lower bounds given by (\ref{WS(n+1)}), 
(\ref{WS(n+2)}) and (\ref{WS(n+3)}) for \( n \in [\![8,15]\!] \). The best lower bound
for each integer is highlighted.

@@

\begin{table}[H]

\label{LowerBoundsWS}
\[
\begin{array}{c}
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-4}
		\cellcolor{yellow}{2-5}
		\cellcolor{yellow}{3-2}
		\cellcolor{yellow}{4-3}
	\Body
		n & 8 & 9 & 10 & 11 \\
		4 \, S(n-1) + 2 & 6\,722 & 21\,146 & 71\,214 & 243\,794 \\
		13 \, S(n-2) + 8 & 6\,976 & 21\,848 & 68\,726 & 231\,447 \\
		42 \, S(n-3) + 24 & 6\,744 & 22\,536 & 70\,584 & 222\,036 \\
	\end{NiceArray}
	\\ \\
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-2}
		\cellcolor{yellow}{2-5}
		\cellcolor{yellow}{3-3}
		\cellcolor{yellow}{4-4}
	\Body
		n & 12 & 13 & 14 & 15 \\
		4 \, S(n-1) + 2 & 815\,314 & 2\,554\,194 & 8\,045\,162 & 27\,061\,154 \\
		13 \, S(n-2) + 8 & 792\,332 & 2\,649\,772 & 8\,301\,132 & 26\,146\,778 \\
		42 \, S(n-3) + 24 & 747\,750 & 2\,559\,840 & 8\,560\,800 &  25\,886\,224 \\
	\end{NiceArray}
\end{array}
\]
\caption{New lower bounds for \( n \in [\![8,15]\!] \)}
\end{table}

@@

With \( S(9) \geqslant 17\,803 \), we found a new lower bound for \(\WS(10)\) using (\ref{WS(n+1)}).
Moreover, (\ref{WS(n+3)}) gives new lower bounds for \(\WS(9)\) and \(WS(14)\).

@@

\qquad In this section, we first gave a new construction which can be seen as an equivalent for weakly sum-free partitions of Abott
and Hanson's construction for sum-free partitions. We then  introduced WSF-templates and generalized this construction. This
allows us to find new lower bounds and new inequalities for weak Schur numbers. One may notice the significant gap
between the former lower bounds for weak Schur numbers obtained by conducting a computer search and the new lower bounds
obtained with WSF-templates (including (\ref{WS(n+1)}) and (\ref{WS(n+2)}).
